Convolutional Neural Networks on Graphs\\ with Fast Localized Spectral Filtering

### 2016-10-29 nips2016/nips2016.tex
nips v6: edition by Xavier (2016-06-29)

%\usepackage{nips_2016}
\usepackage[final]{nips2016}

@@

% Use less space
%\usepackage{titlesec}		% space before / after headings
%\titlespacing\section{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\titlespacing\subsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\usepackage{parskip} \setlength{\parskip}{5pt}
\setlength{\bibsep}{0.2pt}

\title{Convolutional Neural Networks on Graphs\\
with Fast Localized Spectral Filtering}

\author{
  Michaël Defferrard \\ %\thanks{further info} \\
  %Department of Electrical Engineering \\
  %École Polytechnique Fédérale de Lausanne (EPFL) \\
  EPFL, Lausanne, Switzerland \\
  \texttt{michael.defferrard@epfl.ch} \\
  \And % \AND
  Xavier Bresson \\
  EPFL, Lausanne, Switzerland \\
  \texttt{xavier.bresson@epfl.ch} \\
  \And % \AND
  Pierre Vandergheynst \\
  EPFL, Lausanne, Switzerland \\
  \texttt{pierre.vandergheynst@epfl.ch} \\
}

@@

	Convolutional neural networks (CNNs) have greatly improved state-of-the-art
	performances in a number of fields, notably computer vision and natural
	language processing. In this work, we are interested in generalizing the
	formulation of CNNs from low-dimensional regular Euclidean domains, where
	image (2D), video (3D) and audio (1D) are represented, to high-dimensional
	irregular domains represented by graphs. This paper introduces a formulation
	of CNNs on graphs in the context of spectral graph theory. We borrow the
	fundamental tools from the emerging field of signal processing on graphs,
	which provides the necessary mathematical background and efficient numerical
	schemes to design localized graph filters which are fast to evaluate.
	Similarly to classical CNNs, the computational complexity as well as the
	number of parameters to learn are independent of the data dimensionality.
	Numerical experiments on MNIST and 20NEWS demonstrate the ability of the
	model to learn local and stationary features on graphs, as long as the graph
	is well-constructed. This work can be thought as the foundation of a second
	generation of CNNs on graphs built on spectral graph theory, being
	mathematically well defined, bringing better test performance and being
	faster to evaluate and train than the first generation introduced by Bruna,
	LeCun et al.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{images/illustrationCNNgraphs.eps}
%\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.png}
\caption{Architecture of the proposed CNNs on graphs. Notation: $l$ is the
coarsening level, $x^l$ are the downsampled signals at layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
\label{fig:}
\end{figure}

@@

Convolutional neural networks (CNNs) \cite{pro:LeCunBottouBengioHaffner98MNIST}
offer an efficient architecture to extract highly meaningful statistical
patterns in large-scale and high-dimensional datasets. The key success of CNNs
is its advanced ability to learn local stationary structures and compose them to
form multi-scale hierarchical patterns. Precisely, CNNs extract the local
stationarity property of the input data or signals by revealing local features
that are shared across the data domain. These similar features are identified
with localized convolutional filters or kernels, which are learned from the
data. Convolutional filters are shift- or translation-invariant filters, meaning
they are able to recognize identical features independently of their spatial
locations. Localized kernels or compactly supported filters refer to filters
that extract local features independently of the input data size, with a support
size that can be much smaller than the input size.

@@

A generalization to irregular grids is not direct as the building blocks of
CNNs, the convolution and downsampling operators, are not mathematically clear
on these domains. This makes this extension challenging, both theoretically and
implementation-wise. This work leverages the recent works of
\cite{pro:GregorLeCun10LRF, pro:CoatesNg11LRF,
art:BrunaZarembaSzlamLeCun13DLgraphs, art:HenaffBrunaLeCun15DLgraphs,
art:HammondVandergheynstGribonval11GraphWav,
art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to introduce an
efficient implementation of CNNs on irregular domains represented here as
graphs. Graphs are universal representations of heterogeneous pairwise
relationships between possibly high-dimensional data. Graphs can encode complex
geometric structures, and can be studied with strong mathematical tools such as
spectral graph theory, a blend between graph theory and harmonic analysis.

The major bottleneck of generalizing CNNs to graphs, and the primary goal of
this work, is the definition of localized filters which are efficient to
evaluate and learn.
%that are the core elements to extract local stationarity structures.
We will make use of recent tools developed in the context of graph signal
processing (GSP) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
to precisely control the spatial support of graph filters as well as to allow
fast evaluations. Eventually, a generic and efficient implementation of CNNs for
graph signals may pave the way to the development of a potentially large class
of deep learning techniques for graph-based data.

@@

\todo{spatial approaches provide localization by definition}
\todo{why spectral approach?}
While a spatial approach is conceivable, it faces the challenge of matching
local neighborhoods, as shown by \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}.

\todo{Another challenge is the definition of a pooling layer.}
Local stationarity property of data is extracted via localized convolutional
kernels. In this section, we are interested to extract the multi-scale
hierarchical composition property of data. This is efficiently achieved via grid
subsampling or downsampling and pooling, which trades spatial resolution with
feature resolution reducing the learning complexity without compromising the
system performances. In the case of standard CNNs for data lying on regular
grids, the subsampling operation is mathematically sound. The translation
operator commutes with the downsampling operator, hence not affecting the
translation invariance property of the system at different scales of grid
resolutions.

\paragraph{Graph Signal Processing.} GSP is an emerging field that aims at
bridging the gap between signal processing techniques like wavelet analysis and
graph theory such as spectral graph analysis
\cite{art:BelkinNiyogi05LaplaBeltrami, art:VonLuxburg07Tutorial}. A goal is to
generalize fundamental analysis operations for signals from regular grids to
irregular structures embodied by graphs. We refer the reader to
\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
\todo{others for politics} for an
introduction of the field. Standard operations on grids such as convolution,
translation, filtering, dilatation, modulation or downsampling do not extend
directly to graphs and thus require new mathematical definitions while keeping
the original intuitive concepts. In this context, the authors of
\cite{art:HammondVandergheynstGribonval11GraphWav, art:CoifmanLafon06DifMap,
pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet
operators on graphs. In \cite{art:ShumanFarajiVandergheynst16PyramTrans,
art:RamEladCohen11TreeWavelets}, the authors designed a technique to perform
mutli-scale pyramid transforms on graphs. The works of
\cite{pro:TsitsveroBarbarossa15Uncert, pro:PasdeloupAlamiGriponRabbat15Uncert,
art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty principles
on graphs, and showed that intuitive concepts may be lost, but can also produce
enhanced localization principles for signals on graphs. In
\cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown how to
carry out lasso-based signal regularization on graphs, and studied the
intertwined relationships between smoothness and sparsity on graphs. In
\cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors
investigated compressed sensing recovery conditions for graph spectral signal
analysis.

\todo{Model statistical assumptions: locality, stationarity, compositionality.}
These assumptions can be experimentally verified for any dataset. While we know
for them to be true for audio, images and videos, we show experimentally that
they might be satisfied for text documents as long as the graph is properly
constructed.
We also observe that
local stationarity is an assumption that is made about the data, and the CNNs
are fundamentally designed on this data hypothesis. However, we believe this
stationarity assumption to be present in many real-world data, as structured
data are usually composed of repetitive small patterns at different scales of
observation.

\section{Related Work}

@@

Extending CNNs to non-Euclidean domains is a recent line of work. A first
attempt is the so-called local reception fields \cite{pro:GregorLeCun10LRF,
pro:CoatesNg11LRF}, successfully applied to image recognition. The main idea is
to group together features based upon a measure of similarity such as to select
a limited number of connections between two successive layers. While this model
reduces the number of parameters by exploiting the locality assumption, it does
not attempt to exploit any weight-sharing strategy.
%stationarity This approach is however not able to extract similar groups over
%the data domain, as it is not clear how to define convolutional filters on
%these groups.
\todo{read papers}

ShapeNet \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
art:MasciBoscainiBronsteinVandergheynst15ShapeNet} is a generalization of CNNs
to 3D-meshes, a class of smooth low-dimensional non-Euclidean spaces. The
authors used geodesic polar coordinates to define mesh patches and formulated a
deep learning architecture which allows comparison across different manifolds.
They obtained state-of-the-art results for shape recognition tasks.

In \cite{pro:ChenChengMallat14deepHaar} and \cite{pro:RustamovGuibas14deepHaar},
the authors investigated the construction of Haar wavelet transforms on graphs
using a deep hierarchical architecture. They applied the method to object
recognition on sphere, and to sparse reconstruction of faces.
\todo{Learn parameters of a wavelet kernel? Then it is a parametric method,
they impose a shape to the filters.}
\todo{Unknown graph geometry?}

Finally, \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} and
\cite{art:HenaffBrunaLeCun15DLgraphs} introduced a generalization of CNNs on
graphs using graph counterparts of grid convolution and downsampling. They were
able to learn convolutional filters on graphs using the convolution theorem
\cite{book:Mallat99wavelets} that defines convolutions as linear operators that
diagonalize in the Fourier basis (represented by the eigenvectors of the
Laplacian operator). They were thus the first to introduce a spectral
formulation to extend CNNs on graphs. They also introduced a strategy to learn
the graph structure from the data, and applied their model on image recognition,
text categorization and bioinformatics. This approach does however not scale due
to the necessary multiplications by the graph Fourier basis, a matrix of $n^2$
coefficients, where $n$ is the data dimensionality. Despite the cost of
computing this matrix, which requires an EVD on the graph Laplacian, the
dominant cost is the need to multiply the data by this matrix twice (forward and
inverse Fourier transforms) at a cost of $\bO(n^2)$ operations per forward and
backward pass, a computational bottleneck already identified by the authors.
Besides, as they rely on smoothness (via a spline parametrization) in the
Fourier domain to bring localization in the vertex domain, their model does not
provide a precise control over the local support of their kernels, which is
essential to learn localized filters. This work directly builds on theirs, and
we will show how we overcame these limitations and what we further gained.

\subsection{Contributions}

Our contributions can be summed up as follows:
\begin{enumerate}
%\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
\item \textbf{Spectral formulation.} A clean spectral formulation built on
	established tools in GSP.
\item \textbf{Strictly localized filters.} Unlike \cite{\bruna}, the proposed spectral
	filters are strictly localized in a ball of radius $K$, i.e. $K$ hops from
	the central vertex.
\item \textbf{Low computational complexity.} The evaluation of our filters is
	linear w.r.t. the filters support's size $K$ and the number of edges $n \leq
	|\E| \leq n^2$. As most real-world graphs are highly sparse, we have $|\E|
	\ll n^2$ and $|\E| = kn$ for the widespread $k$-neighbors graphs. Moreover,
	this method avoids the Fourier basis altogether, thus the expensive
	eigenvalue decomposition (EVD) necessary to compute it as well as the need
	to store the basis, a matrix of size $n^2$. That is especially relevant when
	working with limited GPU memory. Besides the data, our method only requires
	to store the Laplacian, a sparse matrix of $|\E|$ non-zero values.
\item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
	graphs which, after a rearrangement of the vertices, is analog to pooling of
	1D signals.
\item \textbf{Experimental results.} We present many experiments that ultimately
	show that our formulation is (1) a useful model, (2) computationally
	efficient and (3) superior both in accuracy and complexity to the previous
	generation of spectral graph CNNs introduced in \cite{\bruna}. We also show that
	our formulation performs similarly as a classical CNN on MNIST and study the
	impact of various graph constructions on performance.
\end{enumerate}

Beyond this work, our long-term goal is to make use of GSP tools as much as
possible to formulate mathematically well-designed techniques for CNNs on
graphs. This paper introduces the mathematical and computational foundations of
such models, while future developments may consider to enhance the building
blocks, namely graph filtering and coarsening. In the future, we expect to
further benefit from newly developed tools in GSP to enhance our formulation of
CNNs on graphs.

\section{Learning Fast Localized Spectral Filters} \label{sec:filters}

As there is no notion of spatial translation on graphs, convolutions are
implemented as multiplications in the spectral domain
\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. While a spatial
convolution on an Euclidean domain is localized by definition (via the finite
size of the kernel) and has a linear computational complexity $\bO(n)$, a filter
defined in the spectral domain is not localized by default and costs $\bO(n^2)$
operations because of the necessary multiplication by the graph Fourier basis.
Note that while classical CNNs can be speed-up by leveraging the Fourier basis
(instead of computing the convolutions in the spatial domain)
\cite{mathieu_fast_2013}, it is only advantageous because of the $\bO(n \log(n))$
complexity of the FFT, which has no equivalent on graphs (yet). A good choice of
filter parametrization however allows us to prove that when translated (via a
convolution with a delta) to a given center vertex, the filter is strictly
localized in a ball of radius $K$ around that center vertex. Moreover, a
recursive formulation allows the filtering operation to scale linearly with $n$.

@@

\paragraph{Polynomial parametrization for localized filters.} There are however
two problems with non-parametric filters: (1) they are not localized and (2)
they are of learning complexity $\bO(n)$, the dimensionality of the data. We are
going to solve those two problems together by reducing this complexity to
$\bO(K)$, the support's size of the filter, exactly as it is for standard CNNs.
While \cite{\bruna} solves it by parameterizing $g_\theta(\Lambda)$ as a cubic
B-spline defined by $K$ control points, let us parametrize it as a polynomial
function
\begin{equation} \label{eq:filt_poly}
	g_\theta(\Lambda) = \sum_{k=0}^{K-1} \theta_k \Lambda^k,
\end{equation}
where the parameter $\theta \in \R^K$ is a vector of polynomial coefficients.
The value at vertex $j$ of the filter $g_\theta$ centered at vertex $i$ is given
by $(g_\theta(L) \delta_i)_j = (g_\theta(L))_{i,j} = \sum_k \theta_k
(L^k)_{i,j}$, where the kernel is translated via a convolution with a Kronecker
delta $\delta_i \in \R^n$. By \cite[Lemma
5.2]{art:HammondVandergheynstGribonval11GraphWav}, $d_\G(i,j) > K$ implies
$(L^K)_{i,j} = 0$, where $d_\G$ is the shortest path distance, i.e. the distance
between two vertices is the minimum number of edges in any path connecting them.
Consequently, spectral filters represented by $K^\text{th}$-order polynomials of
the Laplacian are exactly $K$-localized.

% While there is many possible choice of recursive polynomials
\paragraph{Recursive formulation for fast filtering.} While we have shown how to
learn localized filters with $K$ parameters, the cost to filter a signal $x$ as
$y = Ug_\theta(\Lambda)U^Tx$ is still of $\bO(n^2)$ operations because of the
multiplications with the Fourier basis $U$. A way out is to parametrize
$g_\theta(L)$ as a polynomial function that can be computed recursively from
$L$, as $K$ multiplications by a sparse $L$ costs $\bO(K|\E|) \ll \bO(n^2)$. One
such polynomial, traditionally used in GSP to approximate kernels (like
wavelets), is the Chebyshev expansion
\cite{art:HammondVandergheynstGribonval11GraphWav}. Another option, the Lanczos
algorithm \cite{art:SusnjaraPerraudinKressnerVandergheynst15Lanczos}, which
constructs an orthonormal basis of the Krylov subspace $\mathcal{K}_K(L,x) =
\spn\{x,Lx,\ldots,L^{K-1}x\}$, seems attractive because of the coefficients'
independence. It is however more convoluted and thus left as a future work.

@@

MNIST is a dataset of 70,000 digit numbers represented on a 2D grid of size $28
\times 28$, such that data points lie on a space of $784$ dimensions. To test
our model, we constructed a 8-NN (nearest neighbors) graph of the 2D Euclidean
grid, which produced a graph of $n = |\V| = 976$ nodes (784 pixels and 192 fake
nodes) and $|\E| = 3198$ edges.

The LeNet-5-like network architecture and the following hyper-parameters are
borrowed from the TensorFlow MNIST tutorial\footnote{
\url{https://www.tensorflow.org/versions/r0.8/tutorials/mnist/pros}}: dropout
probability of 0.5, regularization weight of $5\times10^{-4}$, initial learning
rate of 0.03, learning rate decay of 0.95, momentum of 0.9. Convolutional layer
have filters of size $5 \times 5$ while graph convolution layers have the same
support of $K = 5^2 = 25$. All models were trained for \todo{20 epochs}.

>>>>

%\usepackage{nips_2016}
%\usepackage[final]{nips_2016}
\usepackage[top=1.2in, bottom=1.2in, left=1.2in, right=1.2in]{geometry} % arxiv

@@

% Use less space
%\usepackage{titlesec}		% space before / after headings
%\titlespacing\section{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\titlespacing\subsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\usepackage{parskip} \setlength{\parskip}{5pt}
%\setlength{\bibsep}{0.2pt} %comment for arxiv

\title{Convolutional Neural Networks on Graphs\\
with Fast Localized Spectral Filtering\\}

%\author{
%  Michaël Defferrard \\ %\thanks{further info} \\
%  %Michael Defferrard \\ %\thanks{further info} \\
%  %Department of Electrical Engineering \\
%  %École Polytechnique Fédérale de Lausanne (EPFL) \\
%  EPFL, Lausanne, Switzerland \\
%  \texttt{michael.defferrard@epfl.ch} \\
%  \And % \AND
%  Xavier Bresson \\
%  EPFL, Lausanne, Switzerland \\
%  \texttt{xavier.bresson@epfl.ch} \\
%  \And % \AND
%  Pierre Vandergheynst \\
%  EPFL, Lausanne, Switzerland \\
%  \texttt{pierre.vandergheynst@epfl.ch} \\
%}

%%Arxiv
\author{
  Micha\"{e}l Defferrard\hspace{1cm}Xavier Bresson\hspace{1cm} Pierre Vandergheynst\\ \\
  Department of Electrical Engineering \\
  EPFL, Lausanne, Switzerland \\
  {\small \texttt{\{michael.defferrard,xavier.bresson,pierre.vandergheynst\}@epfl.ch} } \\\\
}

@@

	Convolutional neural networks (CNNs) have greatly improved state-of-the-art
	performances in a number of fields, notably computer vision and natural
	language processing. In this work, we are interested in generalizing the
	formulation of CNNs from low-dimensional regular Euclidean domains, where
	images (2D), videos (3D) and audios (1D) are represented, to high-dimensional
	irregular domains such as social networks or biological networks represented by graphs. This paper introduces a formulation
	of CNNs on graphs in the context of spectral graph theory. We borrow the
	fundamental tools from the emerging field of signal processing on graphs,
	which provides the necessary mathematical background and efficient numerical
	schemes to design localized graph filters efficient to learn and evaluate. 
	As a matter of fact, we introduce the first technique that offers the same computational complexity than standard CNNs, while being universal to any graph structure. 
	%Similarly to classical CNNs, the computational complexity as well as the
	%number of parameters to learn are independent of the data dimensionality.
	Numerical experiments on MNIST and 20NEWS demonstrate the ability of this novel deep learning system to learn local, stationary, and compositional features on graphs, as long as the graph
	is well-constructed. %This work can be thought as the foundation of a second
	%generation of CNNs on graphs built on spectral graph theory, being
	%mathematically well defined, bringing better test performance and being
	%faster to evaluate and train than the first generation introduced by Bruna,
	%LeCun et al.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{images/illustrationCNNgraphs.eps}
%\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.png}
\caption{Architecture of the proposed CNNs on graphs. Notation: $l$ is the
coarsening level, $x^l$ are the downsampled signals at layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
\label{fig_illustration}
\end{figure}

@@

Convolutional neural networks \cite{pro:LeCunBottouBengioHaffner98MNIST}
offer an efficient architecture to extract highly meaningful statistical
patterns in large-scale and high-dimensional datasets. The key success of CNNs
is its advanced ability to learn local stationary structures and compose them to
form multi-scale hierarchical patterns. Precisely, CNNs extract the local
stationarity property of the input data or signals by revealing local features
that are shared across the data domain. These similar features are identified
with localized convolutional filters or kernels, which are learned from the
data. Convolutional filters are shift- or translation-invariant filters, meaning
they are able to recognize identical features independently of their spatial
locations. Localized kernels or compactly supported filters refer to filters
that extract local features independently of the input data size, with a support
size that can be much smaller than the input size.

@@

A generalization to irregular grids is not straightforward as the building blocks of
CNNs, the convolution, the downsampling, and the pooling operators, cannot be directly applied to these domains. This makes this extension challenging, both theoretically and
implementation-wise. This work leverages the recent works of
\cite{pro:GregorLeCun10LRF, pro:CoatesNg11LRF,
art:BrunaZarembaSzlamLeCun13DLgraphs, art:HenaffBrunaLeCun15DLgraphs,
art:HammondVandergheynstGribonval11GraphWav,
art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to introduce an
efficient implementation of CNNs on irregular domains represented here as
graphs. Graphs are universal representations of heterogeneous pairwise
relationships between possibly high-dimensional data. Graphs can encode complex
geometric structures, and can be studied with strong mathematical tools such as
spectral graph theory \cite{book:Chung97Spectral}, a blend between graph theory and harmonic analysis.

@@

The major bottleneck of generalizing CNNs to graphs, and one of the primary goals of
this work, is the definition of graph localized filters which are efficient to
evaluate and learn.
%that are the core elements to extract local stationarity structures.
We will make use of recent tools developed in the context of graph signal
processing (GSP) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to achieve such goals. Precisely, the main contributions of this work are summarized below:
\begin{enumerate}
%\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
\item \textbf{Spectral formulation.} A graph spectral theoretical formulation of CNNs on graphs built on
	established tools in GSP.
\item \textbf{Strictly localized filters.} Enhancing \cite{\bruna}, the proposed spectral
	filters are provable to be strictly localized in a ball of radius $K$, i.e. $K$ hops from
	the central vertex.  
\item \textbf{Low computational complexity.} The evaluation complexity of our filters is
	linear w.r.t. the filters support's size $K$ and the number of edges $|\E|$. Importantly, as most real-world graphs are highly sparse, we have $|\E| \ll n^2$ and $|\E| = kn$ for the widespread $k$-nearest neighbor (NN) graphs, leading to a linear complexity w.r.t the input data size. Moreover,
	this method avoids the Fourier basis altogether, thus the expensive
	eigenvalue decomposition (EVD) necessary to compute it as well as the need
	to store the basis, a matrix of size $n^2$. That is especially relevant when
	working with limited GPU memory. Besides the data, our method only requires
	to store the Laplacian, a sparse matrix of $|\E|$ non-zero values.
\item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
	graphs which, after a rearrangement of the vertices as a binary tree structure, is analog to pooling of
	1D signals.
\item \textbf{Experimental results.} We present many experiments that ultimately
	show that our formulation is (1) a useful model, (2) computationally
	efficient and (3) superior both in accuracy and complexity to the pioneer spectral graph CNNs introduced in \cite{\bruna}. We also show that
	our graph formulation performs similarly as a classical CNNs on MNIST and study the
	impact of various graph constructions on classification performance.
\end{enumerate}

\noindent
Eventually, generic and efficient implementations of CNNs on
graphs, as proposed here and related works \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF,pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
art:MasciBoscainiBronsteinVandergheynst15ShapeNet,art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs}, is essential to pave the way to the development of a new class of deep learning techniques for graph-based data.

\section{Related Works}
\subsection{Graph Signal Processing} GSP is an emerging field that aims at
bridging the gap between signal processing techniques like wavelet analysis \cite{book:Mallat99wavelets} and
graph theory such as spectral graph analysis
\cite{art:BelkinNiyogi05LaplaBeltrami, art:VonLuxburg07Tutorial}. A goal is to
generalize fundamental analysis operations for signals from regular grids to
irregular structures embodied by graphs. We refer the reader to
\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
 for an
introduction of the field. Standard operations on grids such as convolution,
translation, filtering, dilatation, modulation or downsampling do not extend
directly to graphs and thus require new mathematical definitions while keeping
the original intuitive concepts. In this context, the authors of
\cite{art:HammondVandergheynstGribonval11GraphWav, art:CoifmanLafon06DifMap,
pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet
operators on graphs. In \cite{art:ShumanFarajiVandergheynst16PyramTrans,
art:RamEladCohen11TreeWavelets}, the authors designed a technique to perform
mutli-scale pyramid transforms on graphs. The works of
\cite{pro:TsitsveroBarbarossa15Uncert, pro:PasdeloupAlamiGriponRabbat15Uncert,
art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty principles
on graphs, and showed that intuitive concepts may be lost, but can also produce
enhanced localization principles for signals on graphs. In
\cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown how to
carry out lasso-based signal regularization on graphs, and studied the
intertwined relationships between smoothness and sparsity on graphs. In
\cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors
investigated compressed sensing recovery conditions for graph spectral signal
analysis.

In the future, we expect to
further benefit from newly developed tools in GSP to enhance our mathematical formulation of
CNNs on graphs. This paper introduces the mathematical and computational foundations of
such models, while future developments may consider to enhance the building
blocks, namely graph filtering and coarsening.

\subsection{CNNs on Non-Euclidean Domains}

@@

Extending CNNs to non-Euclidean domains has been a recent line of work. A preliminary work was proposed with 
 the so-called local reception fields \cite{pro:GregorLeCun10LRF,
pro:CoatesNg11LRF}, successfully applied to image recognition. The main idea is
to group together features based upon a measure of similarity such as to select
a limited number of connections between two successive layers. While this model
reduced the number of parameters by exploiting the locality assumption, it did
not attempt to exploit any stationarity property of data, i.e. no weight-sharing strategy. %stationarity This approach is however not able to extract similar groups over
%the data domain, as it is not clear how to define convolutional filters on
%these groups.
%\todo{read papers}

A first attempt to design CNNs on graphs was introduced in \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} by adopting a spatial approach. The idea was to construct a multiresolution graph by exploiting graph neighbourhood structures, and learned a deep neural network. However, such approach did not consider convolution operations, and thus no stationarity property. %By construction, spatial approaches provide filter localization. However, although graph convolution directly in the spatial domain is conceivable, it also faces the challenge of matching local neighborhoods, as shown by \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}.

ShapeNet in \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
art:MasciBoscainiBronsteinVandergheynst15ShapeNet} is a generalization of CNNs
to 3D-meshes, a class of smooth low-dimensional non-Euclidean spaces. The
authors used geodesic polar coordinates to define convolution operations on mesh patches, and formulated a
deep learning architecture which allows comparison across different manifolds.
They obtained state-of-the-art results for 3D shape recognition tasks.

In \cite{pro:ChenChengMallat14deepHaar} and \cite{pro:RustamovGuibas14deepHaar},
the authors investigated the construction of Haar wavelet transforms on graphs
using a deep hierarchical architecture. They applied the method to object
recognition on sphere, and to sparse reconstruction of faces.
%\todo{Learn parameters of a wavelet kernel? Then it is a parametric method,
%they impose a shape to the filters.}
%\todo{Unknown graph geometry?}

Finally, \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} and
\cite{art:HenaffBrunaLeCun15DLgraphs} introduced a generalization of CNNs on
graphs using graph counterparts of grid convolution and downsampling. They were
able to learn convolutional filters on graphs using the convolution theorem
\cite{book:Mallat99wavelets} that defines convolutions as linear operators that
diagonalize in the Fourier basis (represented by the eigenvectors of the
Laplacian operator). They were thus the first to introduce a spectral
formulation to extend CNNs on graphs. They also introduced a strategy to learn
the graph structure from the data, and applied their model on image recognition,
text categorization and bioinformatics. This approach does however not scale up due
to the necessary multiplications by the graph Fourier basis, a matrix of $n^2$
coefficients, where $n$ is the data dimensionality. Despite the cost of
computing this matrix, which requires an EVD on the graph Laplacian, the
dominant cost is the need to multiply the data by this matrix twice (forward and
inverse Fourier transforms) at a cost of $\bO(n^2)$ operations per forward and
backward pass, a computational bottleneck already identified by the authors.
Besides, as they rely on smoothness (via a spline parametrization) in the
Fourier domain to bring localization in the vertex domain, their model does not
provide a precise control over the local support of their kernels, which is
essential to learn localized filters. Our technique leverages on this work, and
we will show how to overcome these limitations and beyond.

\section{Proposed Technique}
Generalizing CNNs to graphs requires three fundamental steps. First step is the design of localized convolutional filters on graphs. Second step is a graph coarsening procedure that groups together similar features on graphs. And the last step is the graph pooling operation that trades spatial resolution for higher filter resolution.

\subsection{Learning Fast Localized Spectral Filters} \label{sec:filters}
The goal of this section is to define fast convolutional localized filters on graphs. There are two strategies to define convolutional localized filters; either from a spatial approach or from a spectral approach. By construction, spatial approaches provide filter localization via the finite size of the kernel. However, although graph convolution directly in the spatial domain is conceivable, it also faces the challenge of matching local neighbourhoods, as pointed out in \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. Consequently, there is no unique mathematical definition of spatial translations on graphs from a spatial perspective. On the other side, a spectral approach provides a well-defined translation operator on graphs via convolutions with a Kronecker delta function implemented in the spectral domain
\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. However, a filter defined in the spectral domain is not naturally localized and translations are costly with $\bO(n^2)$ operations due to multiplications with the graph Fourier basis. Both limitations can be overcome with a special choice of filter parametrization, as explained below.

%Note that while classical CNNs can be speed-up by leveraging the Fourier basis
%(instead of computing the convolutions in the spatial domain)
%\cite{mathieu_fast_2013}, it is only advantageous because of the $\bO(n \log(n))$
%complexity of the FFT, which has no equivalent on graphs (yet).

@@

\paragraph{Polynomial parametrization for localized filters.} There are however
two limitations with non-parametric filters: (1) they are not localized in space and (2)
their learning complexity is in $\bO(n)$, the dimensionality of the data. These issues can be overcome 
with the use of Laplacian-based polynomial spectral filters: 
\begin{equation} \label{eq:filt_poly}
	g_\theta(\Lambda) = \sum_{k=0}^{K-1} \theta_k \Lambda^k,
\end{equation}
where the parameter $\theta \in \R^K$ is a vector of polynomial coefficients.
The value at vertex $j$ of the filter $g_\theta$ centered at vertex $i$ is given
by $(g_\theta(L) \delta_i)_j = (g_\theta(L))_{i,j} = \sum_k \theta_k
(L^k)_{i,j}$, where the kernel is translated via a convolution with a Kronecker
delta function $\delta_i \in \R^n$. By \cite[Lemma
5.2]{art:HammondVandergheynstGribonval11GraphWav}, $d_\G(i,j) > K$ implies
$(L^K)_{i,j} = 0$, where $d_\G$ is the shortest path distance, i.e. the minimum number of edges connecting two vertices on the graph.
Consequently, spectral filters represented by $K^\text{th}$-order polynomials of
the Laplacian are exactly $K$-localized. Besides, their learning complexity is $\bO(K)$, the support size of the filter, and thus the same complexity as in standard CNNs.

% While there is many possible choice of recursive polynomials
\paragraph{Recursive formulation for fast filtering.} While we have shown how to
learn localized filters with $K$ parameters, the cost to filter a signal $x$ as
$y = Ug_\theta(\Lambda)U^Tx$ is still high with $\bO(n^2)$ operations because of the
multiplications with the Fourier basis $U$. A solution of this problem is to parametrize
$g_\theta(L)$ as a polynomial function that can be computed recursively from
$L$, as $K$ multiplications by a sparse $L$ costs $\bO(K|\E|) \ll \bO(n^2)$. One
such polynomial, traditionally used in GSP to approximate kernels (like
wavelets), is the Chebyshev expansion
\cite{art:HammondVandergheynstGribonval11GraphWav}. Another option, the Lanczos
algorithm \cite{art:SusnjaraPerraudinKressnerVandergheynst15Lanczos}, which
constructs an orthonormal basis of the Krylov subspace $\mathcal{K}_K(L,x) =
\spn\{x,Lx,\ldots,L^{K-1}x\}$, seems attractive because of the coefficients'
independence. It is however more convoluted and thus left as a future work.

@@

We recall that MNIST is a dataset of 70,000 digit numbers represented on a 2D grid of size $28
\times 28$, such that data points lie on a space of $784$ dimensions. For
our graph model, we construct a $8$-NN graph of the 2D Euclidean
grid, which produces a graph of $n = |\V| = 976$ nodes (784 pixels and 192 fake
nodes as explained in \secref{pooling}) and $|\E| = 3198$ edges.
